---
title: "Project 1 - Data Syndicate Assignment / Jessie Alwerdt"
author: "Jessie Alwerdt"
date: "7/19/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Data is from https://www.kaggle.com/paultimothymooney/phd-stipends/data# belonging to Paul Mooney

```{r}
getwd()
```

```{r}
setwd("C:/Users/alwer/Documents/Data Syndicate Projects DSS - FB group/Data")
```

#Import dataset

```{r}
library("readxl")
csv <- read_excel("C:/Users/alwer/Documents/Data Syndicate Projects DSS - FB group/Data/CSV_Excel.xlsx")
```



#Check to see what data type each variable is

```{r}
str(csv)
```


```{r}
csv1 <- csv
```


```{r}
csv1$Academic_Year <- as.factor(csv1$Academic_Year)
is.factor(csv$Academic_Year)
levels(csv1$Academic_Year)
```

```{r}
csv1$Program_Year <- as.factor(csv1$Program_Year)
is.factor(csv$Program_Year)
levels(csv1$Program_Year)
```

#8707 Cases

#Missing
#Overall Pay: 21 missing / Min. -900000 / Max. 994000 / Mean 25124
#LW_Ratio: 911 Missing / Min. -34.01 / Max. 40.97 / Mean 1.076
#Academic Year: 4 missing
#Program Year: 1020 missing
#12 M Gross Pay: 2498 missing / Min. 1 / Max. 100000 / Mean = 28240
#9 M Gross Pay: 6233 missing / Min. 5 / Max. 189600 / Mean = 19596
#3 M Gross Pay: 7909 missing / Min. 3 / Max 55816 / Mean = 5043
#Fees: 5355 / Min. 1 / Max. 1000000 / Mean = 2870

```{r}
summary(csv1)
```

```{r}
library(naniar)
vis_miss(csv1)

```


#Detect outliers

```{r}
csv2 <- as.matrix(csv1)
```

```{r}
library(devtools)
#install_github("mdelacre/Routliers")
```

#Outliers for overall pay
#Total of 178 outliers

```{r}
library(Routliers)
res1 <- outliers_mad(x = csv1$Overall_Pay)
res1

plot_outliers_mad(res1, x = csv1$Overall_Pay) 
```

#Outliers for 12_M_Gross_Pay
#Total of 386 Outliers


```{r}
library(Routliers)
res1 <- outliers_mad(x = csv1$Twelve_M_Gross_Pay)
res1
plot_outliers_mad(res1, x = csv1$Twelve_M_Gross_Pay)
```

#Outliers for 9_M_Gross_Pay
#Total of 143 Outliers

```{r}
library(Routliers)
res1 <- outliers_mad(x = csv1$Nine_M_Gross_Pay)
res1
plot_outliers_mad(res1, x = csv1$Nine_M_Gross_Pay)
```

#Had to rename variable due to not be able to reference it beginning with a number

```{r}
library(plyr)
csv1 <- rename(csv1, c("3_M_Gross_Pay" = "Three_M_Gross_Pay"))

```

#Outliers for 3_M_Gross_Pay
#Total of 27 Outliers

```{r}
library(Routliers)
res1 <- outliers_mad(x = csv1$Three_M_Gross_Pay)
res1
plot_outliers_mad(res1, x = csv1$Three_M_Gross_Pay)
```



```{r}
library(dplyr)
table(csv1$Academic_Year, csv1$Program_Year)

```

#Highest year was 2016-2017 (2016 - 2020)

```{r}
library(ggplot2)
table2 <- table(csv1$Academic_Year)
prop.table(table2)
ggplot(csv1, aes(x = csv1$Academic_Year)) +
        geom_bar() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#Drop of cases with each program year

```{r}
table2 <- table(csv1$Program_Year)
prop.table(table2)
ggplot(csv1, aes(x = csv1$Program_Year)) +
        geom_bar() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#Look into the character values

```{r}
library("wordcloud")
library("tm")
wordcloud(words = csv1$University, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```

```{r, warning=FALSE}
wordcloud(words = csv1$Department, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```



#Wanted to do so much more
#Would like to replicate in SAS and Python environment
#Would like to look at cleaning this better
#Would like to look into the missing cases more
#Outliers need to be examined closer. Some are most likely errors. 

